\documentclass[12pt, a4paper]{article}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{cite}
\usepackage{titlesec}
\usepackage{tikz}
\usetikzlibrary{positioning, shapes.geometric, arrows.meta}

\title{Comprehensive Methods for Microservice Decomposition}
\author{FYP Proposal}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This document outlines four highly concrete, distinct methodologies for automated microservice decomposition. Furthermore, \textbf{every single method} has been engineered to strictly fulfill all four primary project objectives (O1: Chunking, O2: Hierarchical Embeddings, O3: Optimization, O4: Tool Development) while maintaining its own unique architectural paradigm.
\end{abstract}

\clearpage
\section{Graph-Contrastive Hierarchical Community Embedding (GC-HCE) Pipeline}

\subsection{Genesis}
\begin{itemize}
    \item \textbf{Sources Consulted:} Citations \cite{jin2021, chaieb2024, bolanowski2022, kalia2020, kalia2021}.
    \item \textbf{Influence:} Citation \cite{jin2021} explicitly suggests clustering the call graph and building hierarchical embeddings. Citations \cite{chaieb2024} and \cite{bolanowski2022} provide the exact metrics (Structural Modularity, Interface Number, ICP) that must be optimized. This method directly translates those theoretical notes into a concrete Graph Neural Network pipeline.
\end{itemize}

\subsection{Methodology Description}
The GC-HCE pipeline is designed to extract microservices by looking at both the macro-structure and micro-semantics of a monolithic codebase. It operates in three main phases:
\begin{enumerate}
    \item \textbf{Graph Chunking via Louvain Community Detection:} First, the monolith's static call graph is analyzed. We apply \textit{Louvain Community Detection}, a network science algorithm that efficiently groups densely connected nodes (classes) together while cutting sparse links. This forms our coarse, high-level ``macro'' communities.
    \item \textbf{Hierarchical Embeddings Generation:} An embedding is a numerical vector representing code. Here, we build a \textit{Hierarchical Embedding} by concatenating two distinct vectors: a macro-level vector (generated by a Graph Neural Network processing the communities) and a micro-level vector (generated by CodeBERT processing individual function semantics). This ensures the model understands both the structural role and the business logic of every component.
    \item \textbf{Contrastive Learning Optimization:} The embeddings are refined using a \textit{Contrastive Multi-Metric Loss}. Contrastive learning is a machine learning technique that forces similar items closer together in vector space and dissimilar items further apart. By injecting architectural metrics (like Structural Modularity and Inter-partition Communication) directly into this loss function, we forcefully decouple the microservices before clustering them via K-Means.
\end{enumerate}

\subsection{Objective Coverage}
\begin{itemize}
    \item \textbf{O1 (Chunking):} Employs the Louvain community detection algorithm on the static call graph to identify initial macro-chunks.
    \item \textbf{O2 (Hierarchical Embeddings):} Concatenates macro-embeddings (GraphSAGE representations of communities) with micro-embeddings (CodeBERT vectors of specific functions).
    \item \textbf{O3 (Enhanced Optimization):} Uses a contrastive loss function explicitly weighting Triplet Loss, Structural Modularity (SM), and Inter-partition Communication (ICP) to heavily penalize cross-boundary method calls.
    \item \textbf{O4 (Tool Development):} Packaged as a GitHub Action that parses a repository on commit and generates architectural boundary recommendations in a pull request.
\end{itemize}

\subsection{Evaluation Metrics}
The primary evaluation of the GC-HCE pipeline will focus on the cohesiveness and independence of the generated microservices.
\begin{itemize}
    \item \textbf{Structural Modularity (SM):} Assesses whether the community detection appropriately groups highly cohesive classes together.
    \item \textbf{Inter-partition Communication (ICP):} Used to penalize and evaluate excessive runtime method calls between the generated macro-micro clusters.
\end{itemize}

\begin{figure}[htbp]
    \centering
    \resizebox{\linewidth}{!}{%
    \begin{tikzpicture}[
        box/.style={draw, rectangle, rounded corners, align=center, minimum width=2.5cm, minimum height=1cm, fill=blue!5},
        arrow/.style={->, >=Stealth, thick}
    ]
    \node[box] (ast) {Monolith AST \& Call Graph};
    \node[box, below left=1cm and -1cm of ast] (louvain) {Louvain Community\\Detection};
    \node[box, below right=1cm and -1cm of ast] (codebert) {Function-level\\CodeBERT Embeddings};
    \node[box, below=1cm of louvain] (macro) {Macro-level\\GraphSAGE};
    \node[box, below right=1cm and -0.5cm of macro] (fusion) {Concatenation Fusion};
    \node[box, below=1cm of fusion] (hier) {Hierarchical Embedding};
    \node[box, below=1cm of hier] (loss) {Contrastive Multi-Metric Loss};
    \node[box, right=1.5cm of loss] (kmeans) {K-Means\\Clustering};
    \node[box, right=1.5cm of kmeans] (services) {Final Microservices};

    \draw[arrow] (ast) -- (louvain);
    \draw[arrow] (ast) -- (codebert);
    \draw[arrow] (louvain) -- (macro);
    \draw[arrow] (macro) |- (fusion);
    \draw[arrow] (codebert) |- (fusion);
    \draw[arrow] (fusion) -- (hier);
    \draw[arrow] (hier) -- (loss);
    \draw[arrow] (loss) -- node[above] {Optimize} (kmeans);
    \draw[arrow] (kmeans) -- (services);
    \draw[arrow, dashed] (loss.west) -- ++(-1,0) |- (fusion.west);
    \end{tikzpicture}%
    }
    \caption{Workflow visualization for the GC-HCE Pipeline.}
    \label{fig:gchce}
\end{figure}

\clearpage
\section{Auto-Encoder Driven Multi-Granularity GNN}

\subsection{Genesis}
\begin{itemize}
    \item \textbf{Sources Consulted:} Citations \cite{mathai2022, lin2023, bolanowski2022, kalia2021}.
    \item \textbf{Influence:} Citation \cite{mathai2022} highlights the prior study's use of simple embeddings and the need to combine GNNs. Citation \cite{bolanowski2022} mentions Reconstruction Loss acting as dimensionality reduction. This method builds a two-stream GNN Auto-Encoder to leverage that exact reconstruction mechanism.
\end{itemize}

\subsection{Methodology Description}
This method leverages generative deep learning principles to compress the codebase's structural complexity into a clean, lower-dimensional space.
\begin{enumerate}
    \item \textbf{Spectral Clustering Initialization:} The dependency matrix of the codebase is passed through \textit{Spectral Clustering}. Unlike standard clustering, which groups data based on geometric proximity, spectral clustering groups items based on graph connectivity by analyzing the matrix's eigenvalues. This identifies winding, complex dependency chains without breaking them.
    \item \textbf{Two-Stream Autoencoder Setup:} An \textit{Autoencoder} is a neural network designed to compress data into a bottleneck and then perfectly reconstruct it, forcing the model to learn only the most essential structural patterns. We use two parallel GNN streams: one for macro-nodes (clusters) and one for micro-nodes (classes).
    \item \textbf{Cross-Attention Fusion:} The two streams are merged using a \textit{Cross-Attention Mechanism}, a neural network layer that allows the micro-stream to selectively focus on relevant context from the macro-stream, and vice versa. The resulting latent space perfectly captures bounded contexts, allowing Agglomerative Clustering to easily identify robust microservice boundaries.
\end{enumerate}

\subsection{Objective Coverage}
\begin{itemize}
    \item \textbf{O1 (Chunking):} Uses Spectral Clustering on the structural dependency matrix to segment the codebase into highly cohesive initial chunks.
    \item \textbf{O2 (Hierarchical Embeddings):} A two-stream GNN architecture where stream one processes chunk-level macro-nodes and stream two processes class-level micro-nodes, fusing them via a cross-attention mechanism.
    \item \textbf{O3 (Enhanced Optimization):} Focuses heavily on Reconstruction loss combined with Clustering loss and Interface Number (IFN) penalties to ensure decoupled microservices.
    \item \textbf{O4 (Tool Development):} Delivered as a standalone Command Line Interface (CLI) tool capable of ingesting monolithic repositories and exporting decoupled, Docker-ready service definitions.
\end{itemize}

\subsection{Evaluation Metrics}
Because this approach heavily relies on latent space representation, the evaluation will measure both architectural quality and structural fidelity.
\begin{itemize}
    \item \textbf{Interface Number (IFN):} Evaluates whether the cross-attention fusion successfully minimizes exposed endpoints between bounded contexts.
    \item \textbf{Reconstruction Error \& Silhouette Score:} Measures how well the microservices form distinct clusters in the latent vector space.
\end{itemize}

\begin{figure}[htbp]
    \centering
    \resizebox{\linewidth}{!}{%
    \begin{tikzpicture}[
        box/.style={draw, rectangle, rounded corners, align=center, minimum width=2.5cm, minimum height=1cm, fill=green!5},
        arrow/.style={->, >=Stealth, thick}
    ]
    \node[box] (dep) {Dependency Matrix};
    \node[box, below=1cm of dep] (spec) {Spectral Clustering\\Chunking};
    \node[box, below left=1cm and -0.5cm of spec] (stream1) {Stream 1:\\Macro-Node GNN};
    \node[box, below right=1cm and -0.5cm of spec] (stream2) {Stream 2:\\Micro-Node GNN};
    \node[box, below right=1cm and -1.5cm of stream1] (fusion) {Cross-Attention Fusion};
    \node[box, below=1cm of fusion] (latent) {Hierarchical Latent Space};
    \node[box, below=1cm of latent] (loss) {Auto-Encoder Loss};
    \node[box, right=1.5cm of loss] (agglo_clust) {Agglomerative\\Clustering};
    \node[box, right=1.5cm of agglo_clust] (services) {Clustered Services};

    \draw[arrow] (dep) -- (spec);
    \draw[arrow] (spec) -- (stream1);
    \draw[arrow] (spec) -- (stream2);
    \draw[arrow] (stream1) |- (fusion);
    \draw[arrow] (stream2) |- (fusion);
    \draw[arrow] (fusion) -- (latent);
    \draw[arrow] (latent) -- (loss);
    \draw[arrow] (loss) -- (agglo_clust);
    \draw[arrow] (agglo_clust) -- (services);
    \draw[arrow, dashed] (loss.west) -- ++(-1,0) |- (fusion.west);
    \end{tikzpicture}%
    }
    \caption{Workflow visualization for the Auto-Encoder Driven GNN.}
    \label{fig:autoencoder}
\end{figure}

\clearpage
\section{LLM-Bootstrapped Reinforcement Learning Decomposition (RL-MD)}

\subsection{Genesis}
\begin{itemize}
    \item \textbf{Sources Consulted:} Citations \cite{mathai2022, mitchell2007, jin2021, kalia2020}.
    \item \textbf{Influence:} Citation \cite{mathai2022} explicitly states that outputs from decompositions can be used by an LLM-based code generation system to iteratively improve its output. This method formalizes that feedback loop using Reinforcement Learning (RL), where the LLM's feedback and architectural metrics (from \cite{mitchell2007} and \cite{jin2021}) act as the reward signal.
\end{itemize}

\subsection{Methodology Description}
Instead of relying on traditional neural network training (backpropagation), this approach treats microservice decomposition as an active, decision-making game.
\begin{enumerate}
    \item \textbf{Agglomerative Code Chunking:} The codebase is initially segmented using an \textit{Agglomerative} (bottom-up) approach. It treats every function as an individual microservice and iteratively merges the most heavily interacting functions together into structural chunks.
    \item \textbf{Semantic Stacking:} The structural graph embeddings are stacked with purely semantic embeddings generated by a Large Language Model (LLM). This gives the system deep insight into the business domain of each chunk.
    \item \textbf{Reinforcement Learning (RL) Boundary Generation:} An \textit{RL Agent} takes control of the architecture. Reinforcement learning is an AI paradigm where an agent learns to make decisions by receiving rewards or penalties. The agent actively shifts classes across microservice boundaries. Its \textit{Reward Function} is purely mathematical: it receives positive points for increasing Structural Modularity (SM) and severe penalties for increasing Interface Number (IFN) or Inter-partition Communication (ICP).
\end{enumerate}

\subsection{Objective Coverage}
\begin{itemize}
    \item \textbf{O1 (Chunking):} Leverages structural modularity (SM) maximization directly during the initial code chunking phase using a greedy agglomerative graph approach.
    \item \textbf{O2 (Hierarchical Embeddings):} Stacks semantic embeddings (generated via an LLM for fine-grained functions) on top of structural embeddings (generated via graph algorithms for coarse communities).
    \item \textbf{O3 (Enhanced Optimization):} Replaces standard backpropagation with a Reinforcement Learning agent. The agent adjusts cluster boundaries to maximize a reward function defined exactly by the multi-metric equation: $Reward = SM - IFN - ICP$.
    \item \textbf{O4 (Tool Development):} Built as an interactive web dashboard that connects to GitHub, visualizes the RL agent's boundary shifting process in real-time, and allows architects to manually approve or reject boundaries.
\end{itemize}

\subsection{Evaluation Metrics}
Since the RL agent directly uses the metrics as its reward function, evaluation focuses on the convergence of these metrics and the human-in-the-loop qualitative feedback.
\begin{itemize}
    \item \textbf{Reward Convergence ($SM - IFN - ICP$):} Measures the agent's ability to maximize structural modularity while simultaneously minimizing interface exposure and inter-partition communication.
    \item \textbf{Architectural Review Approval Rate:} A qualitative metric capturing the percentage of boundary shifts accepted by senior software architects via the interactive dashboard.
\end{itemize}

\begin{figure}[htbp]
    \centering
    \resizebox{\linewidth}{!}{%
    \begin{tikzpicture}[
        box/.style={draw, rectangle, rounded corners, align=center, minimum width=2.5cm, minimum height=1cm, fill=orange!5},
        arrow/.style={->, >=Stealth, thick}
    ]
    \node[box] (agglo) {Agglomerative Code\\Chunking};
    \node[box, right=1.5cm of agglo] (llm) {LLM Semantic\\Embeddings};
    \node[box, below=1cm of agglo] (graph) {Graph Structural\\Embeddings};
    \node[box, below right=1cm and -1.5cm of graph] (stack) {Hierarchical Stack};
    \node[box, below=1cm of stack] (rl) {RL Agent Boundary Generator};
    \node[box, below=1cm of rl] (reward) {Reward Function\\Evaluation};
    \node[box, right=1.5cm of reward] (final_clust) {RL Boundary\\Clustering};
    \node[box, right=1.5cm of final_clust] (web) {Interactive Web Dashboard};

    \draw[arrow] (agglo) -- (graph);
    \draw[arrow] (graph) |- (stack);
    \draw[arrow] (llm) |- (stack);
    \draw[arrow] (stack) -- (rl);
    \draw[arrow] (rl) -- (reward);
    \draw[arrow, dashed] (reward.west) -- ++(-1.5,0) |- node[near start, left] {Feedback (SM, IFN, ICP)} (rl.west);
    \draw[arrow] (reward) -- node[above] {Optimal} (final_clust);
    \draw[arrow] (final_clust) -- (web);
    \end{tikzpicture}%
    }
    \caption{Workflow visualization for the RL-MD approach.}
    \label{fig:rlmd}
\end{figure}

\clearpage
\section{Adaptive Multi-Metric Contrastive Fusion (AMMCF)}

\subsection{Genesis}
\begin{itemize}
    \item \textbf{Sources Consulted:} Citations \cite{jin2021, kalia2020, chaieb2024, weerasinghe2026}.
    \item \textbf{Influence:} Citation \cite{chaieb2024} introduces the Non-Extreme Distribution (NED) metric used to determine the hyperparameter $k$ (microservice count). This method is directly inspired by making that $k$ calculation dynamic during the training loop, rather than static.
\end{itemize}

\subsection{Methodology Description}
The AMMCF method is heavily focused on dynamically adjusting its own parameters during execution, specifically targeting the notoriously difficult problem of choosing the correct number of microservices ($k$).
\begin{enumerate}
    \item \textbf{Call-Graph Pruning \& Girvan-Newman:} The process begins with \textit{Call-Graph Pruning}, where low-weight, infrequent method calls are completely deleted from the graph to remove noise. The remaining robust graph is passed to the \textit{Girvan-Newman algorithm}, a top-down community detection method that repeatedly removes bottleneck edges to reveal core services.
    \item \textbf{Dynamic Weight Fusion:} When combining macro and micro embeddings, a dynamic weighting layer assesses the complexity of each community. If a community is highly complex, the fusion automatically relies more heavily on the fine-grained micro embeddings.
    \item \textbf{Dynamic Loss Scaling via NED:} The training loop actively monitors the \textit{Non-Extreme Distribution (NED)} metric. NED is a specialized architectural metric that heavily penalizes architectures containing ``god classes'' (massive monolithic services) or ``nano-services'' (services containing only one class). By dynamically scaling the loss function based on the NED score, the system automatically hunts for the mathematically optimal microservice count ($k$) before performing hierarchical clustering.
\end{enumerate}

\subsection{Objective Coverage}
\begin{itemize}
    \item \textbf{O1 (Chunking):} Implements dynamic call-graph pruning (removing low-weight edges) followed by Girvan-Newman community detection to create highly robust initial chunks.
    \item \textbf{O2 (Hierarchical Embeddings):} Utilizes a dynamic weighting scheme to fuse macro-community embeddings and micro-class embeddings. If a community is highly complex, the model dynamically relies more on the micro-embeddings.
    \item \textbf{O3 (Enhanced Optimization):} Trains via a unified loss function that dynamically scales Triplet loss (for separation) and SM/IFN/ICP (for architectural validity) based on the current epoch's NED score, automatically hunting for the optimum $k$ value.
    \item \textbf{O4 (Tool Development):} Distributed as a VS Code / IDE Extension that provides real-time, on-the-fly decomposition feedback and visualizations for developers working within legacy monoliths.
\end{itemize}

\subsection{Evaluation Metrics}
The evaluation for AMMCF revolves around its ability to dynamically discover the optimal microservice count (hyperparameter $k$).
\begin{itemize}
    \item \textbf{Non-Extreme Distribution (NED):} The primary metric to evaluate whether the dynamic loss scaling successfully avoids ``god classes'' (mega-services) and ``nano-services''.
    \item \textbf{Overall Decomposition Quality (ODQ):} A holistic evaluation of the resulting architecture against established baseline monolith benchmarks (e.g., JPetStore, AcmeAir).
\end{itemize}

\begin{figure}[htbp]
    \centering
    \resizebox{\linewidth}{!}{%
    \begin{tikzpicture}[
        box/.style={draw, rectangle, rounded corners, align=center, minimum width=2.5cm, minimum height=1cm, fill=red!5},
        arrow/.style={->, >=Stealth, thick}
    ]
    \node[box] (prune) {Call-Graph Pruning};
    \node[box, below=1cm of prune] (girvan) {Girvan-Newman\\Community Detection};
    \node[box, below=1cm of girvan] (macro) {Macro Embeddings};
    \node[box, right=1.5cm of girvan] (micro) {Micro Embeddings};
    \node[box, below right=1cm and -1.5cm of macro] (fusion) {Dynamic Weight Fusion};
    \node[box, below=1cm of fusion] (unified) {Unified Hierarchical Embedding};
    \node[box, below=1cm of unified] (loss) {Dynamic Loss Scaling (NED)};
    \node[box, right=1.2cm of loss] (hier) {Hierarchical\\Clustering (Dynamic $k$)};
    \node[box, right=1.2cm of hier] (ide) {IDE Extension Output};

    \draw[arrow] (prune) -- (girvan);
    \draw[arrow] (girvan) -- (macro);
    \draw[arrow] (macro) |- (fusion);
    \draw[arrow] (micro) |- (fusion);
    \draw[arrow] (fusion) -- (unified);
    \draw[arrow] (unified) -- (loss);
    \draw[arrow] (loss) -- (hier);
    \draw[arrow] (hier) -- node[above] {Clusters} (ide);
    \draw[arrow, dashed] (loss.west) -- ++(-1.5,0) |- (fusion.west);
    \end{tikzpicture}%
    }
    \caption{Workflow visualization for AMMCF.}
    \label{fig:ammcf}
\end{figure}

\clearpage
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
